---
title: "R Notebook"
output: html_notebook
---

# Community belonging
```{r}
library(tidyverse)
library(here)

community_belonging <- read_csv(here("clean_data/community_belonging.csv"))
```
```{r}
community_belonging %>% 
  filter(area == "Scotland", year == 2019, measurement == "Percent",
         walking_distance_to_nearest_greenspace != "All",
         walking_distance_to_nearest_greenspace != "Don't Know") %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(mean_score = mean(belonging_score)) %>% 
  ggplot(aes(walking_distance_to_nearest_greenspace, mean_score)) +
  geom_col()
```
```{r}
community_belonging %>% 
  filter(area == "Scotland", measurement == "Percent",
         gender != "All") %>% 
  group_by(gender, year) %>% 
  summarise(mean_score = mean(belonging_score)) %>% 
  ggplot(aes(year, mean_score, colour = gender)) +
  geom_line()
```



```{r}
community_belonging %>% 
  filter(area == "Scotland", year == 2019, measurement == "Percent",
         gender != "All") %>% 
  ggplot(aes(community_belonging, value, fill = community_belonging)) +
  geom_col() +
  facet_wrap(~gender)
```
# Neighbourhood rating

```{r}
neighbourhood_rating <- read_csv(here("clean_data/neighbourhood_rating.csv"))
neighbourhood_rating
```
```{r}
neighbourhood_rating %>% 
  filter(area == "Scotland", measurement == "Percent",
         !walking_distance_to_nearest_greenspace %in% c("All", "Don't Know")) %>% 
  group_by(walking_distance_to_nearest_greenspace, year) %>% 
  summarise(mean_score = mean(neighbourhood_score)) %>% 
  ggplot(aes(year, mean_score, colour = walking_distance_to_nearest_greenspace)) +
  geom_line()
```
```{r}
neighbourhood_rating %>% 
  filter(area == "Scotland", measurement == "Percent",
         !walking_distance_to_nearest_greenspace %in% c("All", "Don't Know"),
         year == 2019) %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(mean_score = mean(neighbourhood_score)) %>% 
  ggplot(aes(walking_distance_to_nearest_greenspace, mean_score)) +
  geom_col()
```


```{r}
neighbourhood_rating %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(count = n())
```
```{r}
(102/(3708+915))*100
```
# Join the data?
```{r}
community_belonging
neighbourhood_rating
```
```{r}
neighbourhood_rating_id <- neighbourhood_rating %>% 
  mutate(rating = case_when(
    neighbourhood_rating == "Very good" ~ 5,
    neighbourhood_rating == "Fairly good" ~ 4,
    neighbourhood_rating == "No opinion" ~ 3,
    neighbourhood_rating == "Fairly poor" ~ 2,
    neighbourhood_rating == "Very poor" ~ 1
  ),
  measurement = case_when(
    str_detect(measurement, "Lower") ~ "Lower",
    str_detect(measurement, "Upper") ~ "Upper",
    .default = "Percent"
  )) %>% 
  rowwise() %>% 
  mutate(
    id = str_c(feature_code, year, measurement, gender,
             urban_rural_classification, simd_quintiles, type_of_tenure,
             household_type, ethnicity,
             walking_distance_to_nearest_greenspace, rating)
  ) %>% 
  ungroup() %>% 
  select(-neighbourhood_rating) %>% 
  rename("neighbourhood_value" = value) 

community_belonging_id <- community_belonging %>% 
  mutate(rating = case_when(
    community_belonging == "Not at all strongly" ~ 1,
    community_belonging == "Not very strongly" ~ 2,
    community_belonging == "Don't know" ~ 3,
    community_belonging == "Fairly strongly" ~ 4,
    community_belonging == "Very strongly" ~ 5
  ),
  measurement = case_when(
    str_detect(measurement, "Lower") ~ "Lower",
    str_detect(measurement, "Upper") ~ "Upper",
    .default = "Percent"
  )) %>% 
  rowwise() %>% 
  mutate(
    id = str_c(feature_code, year, measurement, gender,
             urban_rural_classification, simd_quintiles, type_of_tenure,
             household_type, ethnicity,
             walking_distance_to_nearest_greenspace, rating)
  ) %>% 
  ungroup() %>% 
  select(-community_belonging) %>% 
  rename("community_value" = value)
```

```{r}
community_belonging_id %>% 
  summarise(across(.cols = everything(), .fns = ~sum(is.na(.x))))

neighbourhood_rating_id %>% 
  summarise(across(.cols = everything(), .fns = ~sum(is.na(.x))))

```

```{r}
ratings_joined <- neighbourhood_rating_id %>% 
  select(-c(feature_code, year, measurement, gender,
            urban_rural_classification, simd_quintiles, type_of_tenure,
            household_type, ethnicity, area,
            walking_distance_to_nearest_greenspace, rating)) %>% 
  full_join(community_belonging_id, by = join_by(id)) %>% 
  select(id, rating, measurement, community_value, belonging_score,
         neighbourhood_value, neighbourhood_score, area, year,
         gender:walking_distance_to_nearest_greenspace) %>% 
  arrange(area, year, rating)
```

The missing values seen in the raw analysis are now causing issues.
It will be easier to do the analysis on the datasets separately. 

```{r}
community_belonging_id %>% 
  filter(feature_code == "S12000036", year == "2016", measurement == "Lower", rating == 3)
```


```{r}
ratings_joined %>% 
    filter(area == "Scotland", measurement == "Percent",
         walking_distance_to_nearest_greenspace != "All") %>% 
  group_by(walking_distance_to_nearest_greenspace, year) %>% 
  summarise(mean_community = mean(belonging_score),
            mean_neighbourhood = mean(neighbourhood_score)) %>% 
  ggplot(aes(year, mean_community, colour = walking_distance_to_nearest_greenspace)) +
  geom_line()

ratings_joined %>% 
    filter(area == "Scotland", measurement == "Percent",
         walking_distance_to_nearest_greenspace != "All") %>% 
  group_by(walking_distance_to_nearest_greenspace, year) %>% 
  summarise(mean_community = mean(belonging_score),
            mean_neighbourhood = mean(neighbourhood_score)) %>% 
  ggplot(aes(year, mean_neighbourhood, colour = walking_distance_to_nearest_greenspace)) +
  geom_line()
```
# Modelling

## Neighbourhood
```{r}
library(ggfortify)
library(GGally)
library(rpart)
library(rpart.plot)
```
```{r}
n_data <- nrow(neighbourhood_rating)
test_prop <- 0.2

test_index <- sample(1:n_data, n_data * test_prop)

neighbour_test <- slice(neighbourhood_rating, test_index)
neighbour_train <- slice(neighbourhood_rating, -test_index)
```

```{r message=FALSE, warning=FALSE}
neighbour_train %>% 
  filter(measurement == "Percent") %>% 
  select(neighbourhood_score, year, gender, urban_rural_classification, simd_quintiles) %>% 
  ggpairs()
  
neighbour_train %>% 
  filter(measurement == "Percent") %>% 
  select(neighbourhood_score, walking_distance_to_nearest_greenspace, ethnicity,
         household_type, type_of_tenure) %>% 
  ggpairs()
```
```{r}
mod1a <- lm(neighbourhood_score ~ urban_rural_classification, neighbour_train)
mod1b <- lm(neighbourhood_score ~ simd_quintiles, neighbour_train)

summary(mod1a)
summary(mod1b)
```
```{r}
autoplot(mod1a)
autoplot(mod1b)
```
```{r}
mod2a <- lm(neighbourhood_score ~ simd_quintiles + urban_rural_classification,
            neighbour_train)
mod2b <- lm(neighbourhood_score ~ year + gender + urban_rural_classification +
    simd_quintiles + walking_distance_to_nearest_greenspace + ethnicity +
         household_type + type_of_tenure,
    neighbour_train)

summary(mod2a)
summary(mod2b)
```
```{r}
mod3a <- lm(neighbourhood_score ~ gender + urban_rural_classification +
    simd_quintiles + ethnicity + household_type + type_of_tenure,
    neighbour_train)
mod3b <- lm(neighbourhood_score ~ gender + urban_rural_classification +
    simd_quintiles + household_type + type_of_tenure,
    neighbour_train)

summary(mod3a)
summary(mod3b)
```
### Conclusions

* Important factors are:
  - gender
  - urban/rural
  - SIMD
  - household type
  - type of tenure
  
Most likely people to give higher neighbourhood ratings:
  - 80% least deprived
  - home owners (outright)
  - rural setting
  - home owners (mortgage)
  - pensioners
  - females

## Community

```{r}
n_data_2 <- nrow(community_belonging)

test_index <- sample(1:n_data_2, n_data_2 * test_prop)

community_test <- slice(community_belonging, test_index)
community_train <- slice(community_belonging, -test_index)
```

**GOT TO HERE**

```{r message=FALSE, warning=FALSE}
train %>% 
  filter(measurement == "Percent") %>% 
  select(belonging_score, year, gender, urban_rural_classification, simd_quintiles) %>% 
  ggpairs()
  
train %>% 
  filter(measurement == "Percent") %>% 
  select(belonging_score, walking_distance_to_nearest_greenspace, ethnicity,
         household_type, type_of_tenure) %>% 
  ggpairs()
```
```{r}
mod1a <- lm(belonging_score ~ year + gender + urban_rural_classification +
    simd_quintiles + walking_distance_to_nearest_greenspace + ethnicity +
         household_type + type_of_tenure, train)
summary(mod1a)
```
```{r}
mod1b <- lm(belonging_score ~ gender + urban_rural_classification +
    simd_quintiles + walking_distance_to_nearest_greenspace + ethnicity +
         household_type + type_of_tenure, train)
summary(mod1b)
```
# Analysis

* What is the average community belonging?
* What is the average neighbourhood score?
* Which groups have a significantly higher score?
* Which groups have a significantly lower score?

```{r}
ratings_joined %>% 
  filter(is.na(neighbourhood_value)) #!!!!

ratings_joined %>% 
  filter(measurement == "Percent") %>% 
  select(rating, community_value, neighbourhood_value) %>% 
  mutate(community_score = community_value * rating,
         neighbourhood_score = neighbourhood_value * rating) %>% 
  summarise(avg_community = mean(community_score),
            avg_neighbour = mean(neighbourhood_score))
```








