---
title: "R Notebook"
output: html_notebook
---

# Community belonging
```{r}
library(tidyverse)
library(here)

community_belonging <- read_csv(here("clean_data/community_belonging.csv"))
```
```{r}
community_belonging %>% 
  filter(area == "Scotland", year == 2019, measurement == "Percent",
         walking_distance_to_nearest_greenspace != "All",
         walking_distance_to_nearest_greenspace != "Don't Know") %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(mean_score = mean(score)) %>% 
  ggplot(aes(walking_distance_to_nearest_greenspace, mean_score)) +
  geom_col()
```
```{r}
community_belonging %>% 
  filter(area == "Scotland", measurement == "Percent",
         gender != "All") %>% 
  group_by(gender, year) %>% 
  summarise(mean_score = mean(score)) %>% 
  ggplot(aes(year, mean_score, colour = gender)) +
  geom_line()
```



```{r}
community_belonging %>% 
  filter(area == "Scotland", year == 2019, measurement == "Percent",
         gender != "All") %>% 
  ggplot(aes(community_belonging, value, fill = community_belonging)) +
  geom_col() +
  facet_wrap(~gender)
```
# Neighbourhood rating
```{r}
neighbourhood_rating <- read_csv(here("clean_data/neighbourhood_rating.csv"))
neighbourhood_rating
```
```{r}
neighbourhood_rating %>% 
  filter(area == "Scotland", measurement == "Percent",
         !walking_distance_to_nearest_greenspace %in% c("All", "Don't Know")) %>% 
  group_by(walking_distance_to_nearest_greenspace, year) %>% 
  summarise(mean_score = mean(score)) %>% 
  ggplot(aes(year, mean_score, colour = walking_distance_to_nearest_greenspace)) +
  geom_line()
```
```{r}
neighbourhood_rating %>% 
  filter(area == "Scotland", measurement == "Percent",
         !walking_distance_to_nearest_greenspace %in% c("All", "Don't Know"),
         year == 2019) %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(mean_score = mean(score)) %>% 
  ggplot(aes(walking_distance_to_nearest_greenspace, mean_score)) +
  geom_col()
```


```{r}
neighbourhood_rating %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(count = n())
```
```{r}
(102/(3708+915))*100
```
# Join the data?
```{r}
community_belonging
neighbourhood_rating
```
```{r}
neighbourhood_rating_id <- neighbourhood_rating %>% 
  mutate(rating = case_when(
    neighbourhood_rating == "Very good" ~ 5,
    neighbourhood_rating == "Fairly good" ~ 4,
    neighbourhood_rating == "No opinion" ~ 3,
    neighbourhood_rating == "Fairly poor" ~ 2,
    neighbourhood_rating == "Very poor" ~ 1
  ),
  measurement = case_when(
    str_detect(measurement, "Lower") ~ "Lower",
    str_detect(measurement, "Upper") ~ "Upper",
    .default = "Percent"
  )) %>% 
  rowwise() %>% 
  mutate(
    id = str_c(feature_code, year, measurement, gender,
               urban_rural_classification, simd_quintiles, type_of_tenure,
               household_type, ethnicity,
               walking_distance_to_nearest_greenspace, rating)
  ) %>% 
  ungroup() %>% 
  select(-neighbourhood_rating) %>% 
  rename("neighbourhood_value" = value) 

community_belonging_id <- community_belonging %>% 
  mutate(rating = case_when(
    community_belonging == "Not at all strongly" ~ 1,
    community_belonging == "Not very strongly" ~ 2,
    community_belonging == "Don't know" ~ 3,
    community_belonging == "Fairly strongly" ~ 4,
    community_belonging == "Very strongly" ~ 5
  ),
  measurement = case_when(
    str_detect(measurement, "Lower") ~ "Lower",
    str_detect(measurement, "Upper") ~ "Upper",
    .default = "Percent"
  )) %>% 
  rowwise() %>% 
  mutate(
    id = str_c(feature_code, year, measurement, gender,
               urban_rural_classification, simd_quintiles, type_of_tenure,
               household_type, ethnicity,
               walking_distance_to_nearest_greenspace, rating)
  ) %>% 
  ungroup() %>% 
  select(-community_belonging) %>% 
  rename("community_value" = value)
```

```{r}
community_belonging_id %>% 
  summarise(across(.cols = everything(), .fns = ~sum(is.na(.x))))

neighbourhood_rating_id %>% 
  summarise(across(.cols = everything(), .fns = ~sum(is.na(.x))))

```

```{r}
ratings_joined <- neighbourhood_rating_id %>% 
  select(-c(feature_code, year, measurement, gender,
            urban_rural_classification, simd_quintiles, type_of_tenure,
            household_type, ethnicity, area,
            walking_distance_to_nearest_greenspace, rating)) %>% 
  full_join(community_belonging_id, by = join_by(id)) %>% 
  select(id, rating, measurement, community_value, score,
         neighbourhood_value, score, area, year,
         gender:walking_distance_to_nearest_greenspace) %>% 
  arrange(area, year, rating)
```

The missing values seen in the raw analysis are now causing issues.
It will be easier to do the analysis on the datasets separately. 

```{r}
community_belonging_id %>% 
  filter(feature_code == "S12000036", year == "2016", measurement == "Lower", rating == 3)
```


```{r}
ratings_joined %>% 
  filter(area == "Scotland", measurement == "Percent",
         walking_distance_to_nearest_greenspace != "All") %>% 
  group_by(walking_distance_to_nearest_greenspace, year) %>% 
  summarise(mean_community = mean(score),
            mean_neighbourhood = mean(score)) %>% 
  ggplot(aes(year, mean_community, colour = walking_distance_to_nearest_greenspace)) +
  geom_line()

ratings_joined %>% 
  filter(area == "Scotland", measurement == "Percent",
         walking_distance_to_nearest_greenspace != "All") %>% 
  group_by(walking_distance_to_nearest_greenspace, year) %>% 
  summarise(mean_community = mean(score),
            mean_neighbourhood = mean(score)) %>% 
  ggplot(aes(year, mean_neighbourhood, colour = walking_distance_to_nearest_greenspace)) +
  geom_line()
```
# Modelling

## Neighbourhood
```{r}
library(ggfortify)
library(GGally)
library(rpart)
library(rpart.plot)
```
```{r}
n_data <- nrow(neighbourhood_rating)
test_prop <- 0.2

test_index <- sample(1:n_data, n_data * test_prop)

neighbour_test <- slice(neighbourhood_rating, test_index)
neighbour_train <- slice(neighbourhood_rating, -test_index)
```

```{r message=FALSE, warning=FALSE}
neighbour_train %>% 
  filter(measurement == "Percent") %>% 
  select(score, year, gender, urban_rural_classification,
         simd_quintiles) %>% 
  ggpairs()

neighbour_train %>% 
  filter(measurement == "Percent") %>% 
  select(score, walking_distance_to_nearest_greenspace, ethnicity,
         household_type, type_of_tenure) %>% 
  ggpairs()
```
```{r}
mod1a <- lm(score ~ urban_rural_classification, neighbour_train)
mod1b <- lm(score ~ simd_quintiles, neighbour_train)

summary(mod1a)
summary(mod1b)
```
```{r}
autoplot(mod1a)
autoplot(mod1b)
```
```{r}
mod2a <- lm(score ~ simd_quintiles + urban_rural_classification,
            neighbour_train)
mod2b <- lm(score ~ year + gender + urban_rural_classification +
              simd_quintiles + walking_distance_to_nearest_greenspace + ethnicity +
              household_type + type_of_tenure,
            neighbour_train)

summary(mod2a)
summary(mod2b)
```
```{r}
mod3a <- lm(score ~ gender + urban_rural_classification +
              simd_quintiles + ethnicity + household_type + type_of_tenure,
            neighbour_train)
mod3b <- lm(score ~ gender + urban_rural_classification +
              simd_quintiles + household_type + type_of_tenure,
            neighbour_train)

summary(mod3a)
summary(mod3b)
```
### Conclusions

* Important factors are:
- gender
- urban/rural
- SIMD
- household type
- type of tenure

Most likely people to give higher neighbourhood ratings:
- 80% least deprived
- home owners (outright)
- rural setting
- home owners (mortgage)
- pensioners
- females

## Community

```{r}
n_data_2 <- nrow(community_belonging)

test_index <- sample(1:n_data_2, n_data_2 * test_prop)

community_test <- slice(community_belonging, test_index)
community_train <- slice(community_belonging, -test_index)
```


```{r message=FALSE, warning=FALSE}
community_train %>% 
  filter(measurement == "Percent") %>% 
  select(score, year, gender, urban_rural_classification, simd_quintiles) %>% 
  ggpairs()

community_train %>% 
  filter(measurement == "Percent") %>% 
  select(score, walking_distance_to_nearest_greenspace, ethnicity,
         household_type, type_of_tenure) %>% 
  ggpairs()
```
```{r}
mod1a <- lm(score ~ year + gender + urban_rural_classification +
              simd_quintiles + walking_distance_to_nearest_greenspace + ethnicity +
              household_type + type_of_tenure, community_train)
summary(mod1a)
```
```{r}
mod1b <- lm(score ~ gender + urban_rural_classification +
              simd_quintiles + walking_distance_to_nearest_greenspace + ethnicity +
              household_type + type_of_tenure, community_train)
summary(mod1b)
```

### Conclusions

* Important factors are:
- type of tenure
- SIMD
- ethnicity
- household type
- urban/rural
- walking distance to greenspace
- gender

*Least* likely people to feel greater community belonging:
- private rented
- non-white ethnicity
- people who don't where where the nearest greenspace is
- other types of tenure

# Analysis

* What is the average community belonging?
* What is the average neighbourhood score?
* Which groups have a significantly higher score?
* Which groups have a significantly lower score?

```{r}
library(infer)
```

## Community/type of tenure
```{r}
average_community_belonging <- community_belonging %>% 
  filter(measurement == "Percent") %>% 
  summarise(avg_rating = mean(score)) %>% 
  pull()

average_community_belonging
```

```{r}
community_belonging %>% 
  filter(measurement == "Percent") %>% 
  group_by(type_of_tenure) %>% 
  summarise(avg_rating = mean(score),
            overall_avg = average_community_belonging)
```
### H0: `Owned Mortgage/Loan` and `Owned Outright` have the same mean (and can therefore be grouped)

```{r}
vars <- c("Owned Outright", "Owned Mortgage/Loan")

null_dist <- community_belonging %>% 
  filter(type_of_tenure %in% vars) %>% 
  specify(score ~ type_of_tenure) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 500, type = "permute") %>% 
  calculate(stat = "diff in means",
            order = vars)

obs_stat <- community_belonging %>% 
  filter(type_of_tenure %in% vars) %>% 
  specify(score ~ type_of_tenure) %>% 
  calculate(stat = "diff in means",
            order = vars)
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "both")

```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "both")
```
At 95% confidence level I cannot be confident that they are different. I will therefore leave them as separate groups.

## H0: `Owned Outright` mean is greater than the average mean

```{r}
null_dist <- community_belonging %>% 
  filter(type_of_tenure == "Owned Outright") %>% 
  #select(score) %>% 
  specify(response = score) %>% 
  hypothesise(null = "point",
              mu = average_community_belonging) %>% 
  generate(reps = 500, type = "bootstrap") %>% 
  calculate(stat = "mean")

obs_stat <- community_belonging %>% 
  filter(type_of_tenure == "Owned Outright") %>%
  specify(response = score) %>% 
  calculate(stat = "mean")
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "right")
```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "both")
```

```{r}
null_dist <- community_belonging %>% 
  select(score, type_of_tenure) %>% 
  mutate(owned_outright = (type_of_tenure == "Owned Outright")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ owned_outright) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 500, type = "permute") %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))

obs_stat <- community_belonging %>% 
  select(score, type_of_tenure) %>% 
  mutate(owned_outright = (type_of_tenure == "Owned Outright")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ owned_outright) %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "right")
```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "both")
```
### Owned outright has a statistically higher community rating than the average

## `Owned Mortgage/Loan`

```{r}
null_dist <- community_belonging %>% 
  select(score, type_of_tenure) %>% 
  mutate(owned_mortgage = (type_of_tenure == "Owned Mortgage/Loan")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ owned_mortgage) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 500, type = "permute") %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))

obs_stat <- community_belonging %>% 
  select(score, type_of_tenure) %>% 
  mutate(owned_mortgage = (type_of_tenure == "Owned Mortgage/Loan")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ owned_mortgage) %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "right")
```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "right")
```
### Owned outright has a statistically higher community rating than the average

## `Private Rented`

```{r}
null_dist <- community_belonging %>% 
  select(score, type_of_tenure) %>% 
  mutate(rented = (type_of_tenure == "Private Rented")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ rented) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 500, type = "permute") %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))

obs_stat <- community_belonging %>% 
  select(score, type_of_tenure) %>% 
  mutate(rented = (type_of_tenure == "Private Rented")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ rented) %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "left")
```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "left")
```
### Private rented has a statistically lower community rating than the average

## `Social Rented`

```{r}
null_dist <- community_belonging %>% 
  filter(type_of_tenure %in% c("All", "Social Rented")) %>% 
  select(score, type_of_tenure) %>% 
  mutate(rented = (type_of_tenure == "Social Rented")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ rented) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 500, type = "permute") %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))

obs_stat <- community_belonging %>% 
  filter(type_of_tenure %in% c("All", "Social Rented")) %>% 
  select(score, type_of_tenure) %>% 
  mutate(rented = (type_of_tenure == "Social Rented")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ rented) %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "both")
```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "both") %>% 
  pull()
```
### Social rented has no impact


# Function for statistical testing
```{r}
test_diff_in_means <- function(data, column, category, n_reps = 500){
  null_dist <- data %>% 
    filter({{column}} %in% c("All", category)) %>% 
    mutate(test = ({{column}} == category)) %>% 
    specify(score ~ test) %>% 
    hypothesise(null = "independence") %>% 
    generate(reps = n_reps, type = "permute") %>% 
    calculate(stat = "diff in means", order = c(TRUE, FALSE))
  
  obs_stat <- data %>% 
    filter({{column}} %in% c("All", category)) %>% 
    mutate(test = ({{column}} == category)) %>% 
    specify(score ~ test) %>%  
    calculate(stat = "diff in means", order = c(TRUE, FALSE))
  
  p <- null_dist %>% 
    get_p_value(obs_stat = obs_stat, direction = "both") %>% 
    pull()
  
  cat(category, ": p = ", p, sep = "")
}
```

```{r}
visualise_diff_in_means <- function(data, column, category, n_reps = 500){
  null_dist <- data %>% 
    filter({{column}} %in% c("All", category)) %>% 
    mutate(test = ({{column}} == category)) %>% 
    specify(score ~ test) %>% 
    hypothesise(null = "independence") %>% 
    generate(reps = n_reps, type = "permute") %>% 
    calculate(stat = "diff in means", order = c(TRUE, FALSE))
  
  obs_stat <- data %>% 
    filter({{column}} %in% c("All", category)) %>% 
    mutate(test = ({{column}} == category)) %>% 
    specify(score ~ test) %>%  
    calculate(stat = "diff in means", order = c(TRUE, FALSE))
  
  null_dist %>% 
    visualise() +
    shade_p_value(obs_stat = obs_stat, direction = "both")
}
```

```{r}
test_diff_in_means(community_belonging, type_of_tenure, "Owned Outright")
visualise_diff_in_means(community_belonging, type_of_tenure, "Owned Outright")
```

# Visualisation of 2019 data for introduction
```{r}
create_2019_community_graph <- function(category){
  community_belonging %>% 
  filter(area == "Scotland", measurement == "Percent", {{category}} != "All",
         year == "2019") %>% 
  select({{category}}, community_belonging, value) %>% 
  ggplot(aes(community_belonging, value, fill = {{category}})) +
  geom_col(position = "dodge") +
  geom_text(aes(label = paste(value, "%")), vjust = -1, position = position_dodge(0.9)) +
  #facet_wrap(~{{category}}) +
  scale_y_continuous(limits = c(0, 50)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        text = element_text(size = 16)) +
  labs(
    title = "Community Belonging in 2019",
    x = "Community Belonging",
    y = "Percentage of survey responses",
    fill = "Classification"
  ) #+
  #scale_fill_manual(values = c("Rural" = "darkgreen", "Urban" = "darkgrey"))
}
```

```{r}
categories <- c("ethnicity", "type_of_tenure", "simd_quintiles", "household_type",
                "urban_rural_classification", "gender", "walking_distance_to_nearest_greenspace")

create_2019_community_graph(ethnicity)
create_2019_community_graph(type_of_tenure)
create_2019_community_graph(simd_quintiles)
create_2019_community_graph(household_type)
create_2019_community_graph(urban_rural_classification)
create_2019_community_graph(gender)
create_2019_community_graph(walking_distance_to_nearest_greenspace)
```

```{r}

```

