---
title: "R Notebook"
output: html_notebook
---

# Community belonging
```{r}
library(tidyverse)
library(here)

community_belonging <- read_csv(here("clean_data/community_belonging.csv")) %>% 
  mutate(community_belonging = factor(community_belonging, 
                                      levels = c("Not at all strongly", 
                                                 "Not very strongly",
                                                 "Don't know",
                                                 "Fairly strongly",
                                                 "Very strongly")))
```
```{r}
community_belonging %>% 
  filter(area == "Scotland", year == 2019, measurement == "Percent",
         walking_distance_to_nearest_greenspace != "All",
         walking_distance_to_nearest_greenspace != "Don't Know") %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(mean_score = mean(score)) %>% 
  ggplot(aes(walking_distance_to_nearest_greenspace, mean_score)) +
  geom_col()
```
```{r}
community_belonging %>% 
  filter(area == "Scotland", measurement == "Percent",
         gender != "All") %>% 
  group_by(gender, year) %>% 
  summarise(total_score = sum(score), total = n()) %>% 
  ggplot(aes(year, total_score, colour = gender)) +
  geom_line()
```



```{r}
community_belonging %>% 
  filter(area == "Scotland", year == 2019, measurement == "Percent",
         gender != "All") %>% 
  ggplot(aes(community_belonging, value, fill = community_belonging)) +
  geom_col() +
  facet_wrap(~gender)
```
# Neighbourhood rating
```{r}
neighbourhood_rating <- read_csv(here("clean_data/neighbourhood_rating.csv")) %>% 
  mutate(neighbourhood_rating = factor(neighbourhood_rating, 
                                       levels = c("Very poor", 
                                                  "Fairly poor",
                                                  "No opinion",
                                                  "Fairly good",
                                                  "Very good"))
  )
neighbourhood_rating
```
```{r}
neighbourhood_rating %>% 
  filter(area == "Scotland", measurement == "Percent",
         walking_distance_to_nearest_greenspace != "All", year == 2019) %>% 
  select(neighbourhood_rating, walking_distance_to_nearest_greenspace, value, score)
```


```{r}
neighbourhood_rating %>% 
  filter(area == "Scotland", measurement == "Percent",
         !walking_distance_to_nearest_greenspace %in% c("All", "Don't Know")) %>% 
  group_by(walking_distance_to_nearest_greenspace, year) %>% 
  summarise(total_score = sum(score)) %>% 
  ggplot(aes(year, total_score, colour = walking_distance_to_nearest_greenspace)) +
  geom_line()
```
```{r}
neighbourhood_rating %>% 
  filter(area == "Scotland", measurement == "Percent",
         !walking_distance_to_nearest_greenspace %in% c("All", "Don't Know"),
         year == 2019) %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(total_score = sum(score)) %>% 
  ggplot(aes(walking_distance_to_nearest_greenspace, total_score)) +
  geom_col()
```


```{r}
neighbourhood_rating %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(count = n())
```
```{r}
(102/(3708+915))*100
```

```{r}
neighbourhood_rating %>% 
  filter(area == "Na h-Eileanan Siar", year == 2016, measurement == "Percent",
         type_of_tenure != "All")
```
```{r}
neighbourhood_rating %>% 
      filter(area == "Na h-Eileanan Siar", measurement == "Percent",
             type_of_tenure != "All") %>% 
      summarise(mean_score = sum(score),
                .by = c(year, type_of_tenure)) %>% 
      ggplot(aes(year, mean_score, colour = type_of_tenure)) +
      geom_line() +
      geom_point(size = 5)
```



# Join the data?
```{r eval=FALSE}
community_belonging
neighbourhood_rating
```
```{r eval=FALSE}
neighbourhood_rating_id <- neighbourhood_rating %>% 
  mutate(rating = case_when(
    neighbourhood_rating == "Very good" ~ 5,
    neighbourhood_rating == "Fairly good" ~ 4,
    neighbourhood_rating == "No opinion" ~ 3,
    neighbourhood_rating == "Fairly poor" ~ 2,
    neighbourhood_rating == "Very poor" ~ 1
  ),
  measurement = case_when(
    str_detect(measurement, "Lower") ~ "Lower",
    str_detect(measurement, "Upper") ~ "Upper",
    .default = "Percent"
  )) %>% 
  rowwise() %>% 
  mutate(
    id = str_c(feature_code, year, measurement, gender,
               urban_rural_classification, simd_quintiles, type_of_tenure,
               household_type, ethnicity,
               walking_distance_to_nearest_greenspace, rating)
  ) %>% 
  ungroup() %>% 
  select(-neighbourhood_rating) %>% 
  rename("neighbourhood_value" = value) 

community_belonging_id <- community_belonging %>% 
  mutate(rating = case_when(
    community_belonging == "Not at all strongly" ~ 1,
    community_belonging == "Not very strongly" ~ 2,
    community_belonging == "Don't know" ~ 3,
    community_belonging == "Fairly strongly" ~ 4,
    community_belonging == "Very strongly" ~ 5
  ),
  measurement = case_when(
    str_detect(measurement, "Lower") ~ "Lower",
    str_detect(measurement, "Upper") ~ "Upper",
    .default = "Percent"
  )) %>% 
  rowwise() %>% 
  mutate(
    id = str_c(feature_code, year, measurement, gender,
               urban_rural_classification, simd_quintiles, type_of_tenure,
               household_type, ethnicity,
               walking_distance_to_nearest_greenspace, rating)
  ) %>% 
  ungroup() %>% 
  select(-community_belonging) %>% 
  rename("community_value" = value)
```

```{r eval=FALSE}
community_belonging_id %>% 
  summarise(across(.cols = everything(), .fns = ~sum(is.na(.x))))

neighbourhood_rating_id %>% 
  summarise(across(.cols = everything(), .fns = ~sum(is.na(.x))))

```

```{r eval=FALSE}
ratings_joined <- neighbourhood_rating_id %>% 
  select(-c(feature_code, year, measurement, gender,
            urban_rural_classification, simd_quintiles, type_of_tenure,
            household_type, ethnicity, area,
            walking_distance_to_nearest_greenspace, rating)) %>% 
  full_join(community_belonging_id, by = join_by(id)) %>% 
  select(id, rating, measurement, community_value, score,
         neighbourhood_value, score, area, year,
         gender:walking_distance_to_nearest_greenspace) %>% 
  arrange(area, year, rating)
```

The missing values seen in the raw analysis are now causing issues.
It will be easier to do the analysis on the datasets separately. 

```{r eval-FALSE}
community_belonging_id %>% 
  filter(feature_code == "S12000036", year == "2016", measurement == "Lower", rating == 3)
```


```{r eval=FALSE}
ratings_joined %>% 
  filter(area == "Scotland", measurement == "Percent",
         walking_distance_to_nearest_greenspace != "All") %>% 
  group_by(walking_distance_to_nearest_greenspace, year) %>% 
  summarise(mean_community = mean(score),
            mean_neighbourhood = mean(score)) %>% 
  ggplot(aes(year, mean_community, colour = walking_distance_to_nearest_greenspace)) +
  geom_line()

ratings_joined %>% 
  filter(area == "Scotland", measurement == "Percent",
         walking_distance_to_nearest_greenspace != "All") %>% 
  group_by(walking_distance_to_nearest_greenspace, year) %>% 
  summarise(mean_community = mean(score),
            mean_neighbourhood = mean(score)) %>% 
  ggplot(aes(year, mean_neighbourhood, colour = walking_distance_to_nearest_greenspace)) +
  geom_line()
```
# Modelling

## Neighbourhood
```{r}
library(ggfortify)
library(GGally)
library(rpart)
library(rpart.plot)
```

## Aggregated modelling
```{r}
aggregated_ratings <- neighbourhood_rating %>% 
  filter(measurement == "Percent", area == "Scotland") %>% 
  group_by(year, gender, urban_rural_classification, 
           simd_quintiles, type_of_tenure, household_type,
           ethnicity, walking_distance_to_nearest_greenspace) %>% 
  summarise(score = sum(score))
```
```{r}
lm(score ~ ., aggregated_ratings) %>% 
  summary()
```
```{r}
aggregated_ratings_gender <- aggregated_ratings %>% 
  filter(gender != "All") 

lm(score ~ gender, aggregated_ratings_gender) %>% 
  summary()
```
```{r}
aggregated_ratings_rural <- aggregated_ratings %>% 
  filter(urban_rural_classification != "All") 
aggregated_ratings_rural
lm(score ~ urban_rural_classification, aggregated_ratings_rural) %>% 
  summary()
```

```{r}
n_data <- nrow(neighbourhood_rating)
test_prop <- 0.2

test_index <- sample(1:n_data, n_data * test_prop)

neighbour_test <- slice(neighbourhood_rating, test_index)
neighbour_train <- slice(neighbourhood_rating, -test_index)
```

```{r message=FALSE, warning=FALSE}
neighbour_train %>% 
  filter(measurement == "Percent") %>% 
  select(score, year, gender, urban_rural_classification,
         simd_quintiles) %>% 
  ggpairs()

neighbour_train %>% 
  filter(measurement == "Percent") %>% 
  select(score, walking_distance_to_nearest_greenspace, ethnicity,
         household_type, type_of_tenure) %>% 
  ggpairs()
```

# Non-aggregated modelling
```{r}
mod1a <- lm(score ~ urban_rural_classification, neighbour_train)
mod1b <- lm(score ~ simd_quintiles, neighbour_train)

summary(mod1a)
summary(mod1b)
```
```{r}
autoplot(mod1a)
autoplot(mod1b)
```
```{r}
mod2a <- lm(score ~ simd_quintiles + urban_rural_classification,
            neighbour_train)
mod2b <- lm(score ~ year + gender + urban_rural_classification +
              simd_quintiles + walking_distance_to_nearest_greenspace + ethnicity +
              household_type + type_of_tenure,
            neighbour_train)

summary(mod2a)
summary(mod2b)
```
```{r}
autoplot(mod2b)
```

```{r}
mod3a <- lm(score ~ gender + urban_rural_classification +
              simd_quintiles + ethnicity + household_type + type_of_tenure,
            neighbour_train)
mod3b <- lm(score ~ gender + urban_rural_classification +
              simd_quintiles + household_type + type_of_tenure,
            neighbour_train)

summary(mod3a)
summary(mod3b)
```

```{r}
autoplot(mod3b)
```

### Conclusions

* Important factors are:
- gender
- urban/rural
- SIMD
- household type
- type of tenure

Most likely people to give higher neighbourhood ratings:
- 80% least deprived
- home owners (outright)
- rural setting
- home owners (mortgage)
- pensioners
- females

## Community

```{r}
n_data_2 <- nrow(community_belonging)

test_index <- sample(1:n_data_2, n_data_2 * test_prop)

community_test <- slice(community_belonging, test_index)
community_train <- slice(community_belonging, -test_index)
```


```{r message=FALSE, warning=FALSE}
community_train %>% 
  filter(measurement == "Percent") %>% 
  select(score, year, gender, urban_rural_classification, simd_quintiles) %>% 
  ggpairs()

community_train %>% 
  filter(measurement == "Percent") %>% 
  select(score, walking_distance_to_nearest_greenspace, ethnicity,
         household_type, type_of_tenure) %>% 
  ggpairs()
```
```{r}
mod1a <- lm(score ~ year + gender + urban_rural_classification +
              simd_quintiles + walking_distance_to_nearest_greenspace + ethnicity +
              household_type + type_of_tenure, community_train)
summary(mod1a)
```
```{r}
mod1b <- lm(score ~ gender + urban_rural_classification +
              simd_quintiles + walking_distance_to_nearest_greenspace + ethnicity +
              household_type + type_of_tenure, community_train)
summary(mod1b)
```

### Conclusions

* Important factors are:
- type of tenure
- SIMD
- ethnicity
- household type
- urban/rural
- walking distance to greenspace
- gender

*Least* likely people to feel greater community belonging:
- private rented
- non-white ethnicity
- people who don't where where the nearest greenspace is
- other types of tenure

# Analysis

* What is the average community belonging?
* What is the average neighbourhood score?
* Which groups have a significantly higher score?
* Which groups have a significantly lower score?

```{r}
library(infer)
```

## Community/type of tenure
```{r}
average_community_belonging <- community_belonging %>% 
  filter(measurement == "Percent") %>% 
  summarise(avg_rating = mean(score)) %>% 
  pull()

average_community_belonging
```

```{r}
community_belonging %>% 
  filter(measurement == "Percent") %>% 
  group_by(type_of_tenure) %>% 
  summarise(avg_rating = mean(score),
            overall_avg = average_community_belonging)
```
### H0: `Owned Mortgage/Loan` and `Owned Outright` have the same mean (and can therefore be grouped)

```{r}
vars <- c("Owned Outright", "Owned Mortgage/Loan")

null_dist <- community_belonging %>% 
  filter(type_of_tenure %in% vars) %>% 
  specify(score ~ type_of_tenure) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 500, type = "permute") %>% 
  calculate(stat = "diff in means",
            order = vars)

obs_stat <- community_belonging %>% 
  filter(type_of_tenure %in% vars) %>% 
  specify(score ~ type_of_tenure) %>% 
  calculate(stat = "diff in means",
            order = vars)
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "both")

```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "both")
```
At 95% confidence level I cannot be confident that they are different. I will therefore leave them as separate groups.

## H0: `Owned Outright` mean is greater than the average mean

```{r}
null_dist <- community_belonging %>% 
  filter(type_of_tenure == "Owned Outright") %>% 
  #select(score) %>% 
  specify(response = score) %>% 
  hypothesise(null = "point",
              mu = average_community_belonging) %>% 
  generate(reps = 500, type = "bootstrap") %>% 
  calculate(stat = "mean")

obs_stat <- community_belonging %>% 
  filter(type_of_tenure == "Owned Outright") %>%
  specify(response = score) %>% 
  calculate(stat = "mean")
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "right")
```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "both")
```

```{r}
null_dist <- community_belonging %>% 
  select(score, type_of_tenure) %>% 
  mutate(owned_outright = (type_of_tenure == "Owned Outright")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ owned_outright) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 500, type = "permute") %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))

obs_stat <- community_belonging %>% 
  select(score, type_of_tenure) %>% 
  mutate(owned_outright = (type_of_tenure == "Owned Outright")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ owned_outright) %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "right")
```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "both")
```
### Owned outright has a statistically higher community rating than the average

## `Owned Mortgage/Loan`

```{r}
null_dist <- community_belonging %>% 
  select(score, type_of_tenure) %>% 
  mutate(owned_mortgage = (type_of_tenure == "Owned Mortgage/Loan")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ owned_mortgage) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 500, type = "permute") %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))

obs_stat <- community_belonging %>% 
  select(score, type_of_tenure) %>% 
  mutate(owned_mortgage = (type_of_tenure == "Owned Mortgage/Loan")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ owned_mortgage) %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "right")
```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "right")
```
### Owned outright has a statistically higher community rating than the average

## `Private Rented`

```{r}
null_dist <- community_belonging %>% 
  filter(type_of_tenure != "All") %>% 
  select(score, type_of_tenure) %>% 
  mutate(rented = (type_of_tenure == "Private Rented")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ rented) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 500, type = "permute") %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))

obs_stat <- community_belonging %>%
  filter(type_of_tenure != "All") %>% 
  select(score, type_of_tenure) %>% 
  mutate(rented = (type_of_tenure == "Private Rented")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ rented) %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "left")
```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "left")
```
### Private rented has a statistically lower community rating than the average

## `Social Rented`

```{r}
null_dist <- community_belonging %>% 
  filter(type_of_tenure %in% c("All", "Social Rented")) %>% 
  select(score, type_of_tenure) %>% 
  mutate(rented = (type_of_tenure == "Social Rented")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ rented) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 500, type = "permute") %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))

obs_stat <- community_belonging %>% 
  filter(type_of_tenure %in% c("All", "Social Rented")) %>% 
  select(score, type_of_tenure) %>% 
  mutate(rented = (type_of_tenure == "Social Rented")) %>% 
  select(-type_of_tenure) %>% 
  specify(score ~ rented) %>% 
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
```


```{r}
null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = "both")
```
```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = "both") %>% 
  pull()
```
### Social rented has no impact


# Function for statistical testing
```{r}
test_diff_in_means <- function(data, column, category, n_reps = 500){
  null_dist <- data %>% 
    filter({{column}} %in% c("All", category)) %>% 
    mutate(test = ({{column}} == category)) %>% 
    specify(score ~ test) %>% 
    hypothesise(null = "independence") %>% 
    generate(reps = n_reps, type = "permute") %>% 
    calculate(stat = "diff in means", order = c(TRUE, FALSE))
  
  obs_stat <- data %>% 
    filter({{column}} %in% c("All", category)) %>% 
    mutate(test = ({{column}} == category)) %>% 
    specify(score ~ test) %>%  
    calculate(stat = "diff in means", order = c(TRUE, FALSE))
  
  p <- null_dist %>% 
    get_p_value(obs_stat = obs_stat, direction = "both") %>% 
    pull()
  
  cat(category, ": p = ", p, sep = "")
}
```

```{r}
visualise_diff_in_means <- function(data, column, category, n_reps = 500){
  null_dist <- data %>% 
    filter({{column}} %in% c("All", category)) %>% 
    mutate(test = ({{column}} == category)) %>% 
    specify(score ~ test) %>% 
    hypothesise(null = "independence") %>% 
    generate(reps = n_reps, type = "permute") %>% 
    calculate(stat = "diff in means", order = c(TRUE, FALSE))
  
  obs_stat <- data %>% 
    filter({{column}} %in% c("All", category)) %>% 
    mutate(test = ({{column}} == category)) %>% 
    specify(score ~ test) %>%  
    calculate(stat = "diff in means", order = c(TRUE, FALSE))
  
  null_dist %>% 
    visualise() +
    shade_p_value(obs_stat = obs_stat, direction = "both")
}
```

```{r}
test_diff_in_means(community_belonging, type_of_tenure, "Owned Outright")
visualise_diff_in_means(community_belonging, type_of_tenure, "Owned Outright")
```

# Visualisation of 2019 data for introduction
```{r}
create_2019_community_graph <- function(category){
  community_belonging %>% 
    filter(area == "Scotland", measurement == "Percent",
           {{category}} != "All",year == "2019") %>% 
    select({{category}}, community_belonging, value) %>% 
    ggplot(aes(community_belonging, value
               , fill = {{category}})) +
    geom_col(position = "dodge") +
    geom_text(aes(label = paste(value, "%")),
              vjust = -1, position = position_dodge(0.9)) +
    scale_y_continuous(limits = c(0, 50)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          text = element_text(size = 16)) +
    labs(
      title = "Community Belonging in 2019",
      x = "Community Belonging",
      y = "Percentage of survey responses",
      fill = "Classification"
    ) 
}
```

```{r}
categories <- c("ethnicity", "type_of_tenure", "simd_quintiles", "household_type",
                "urban_rural_classification", "gender", "walking_distance_to_nearest_greenspace")

create_2019_community_graph(ethnicity)
create_2019_community_graph(type_of_tenure)
create_2019_community_graph(simd_quintiles)
create_2019_community_graph(household_type)
create_2019_community_graph(urban_rural_classification)
create_2019_community_graph(gender)
create_2019_community_graph(walking_distance_to_nearest_greenspace)
#ggsave("graphs_and_images/community_and_greenspace.png")
```

```{r}
create_2019_neighbourhood_graph <- function(category){
  neighbourhood_rating %>% 
    filter(area == "Scotland", measurement == "Percent", {{category}} != "All",
           year == "2019") %>% 
    select({{category}}, neighbourhood_rating, value) %>% 
    ggplot(aes(neighbourhood_rating, value, fill = {{category}})) +
    geom_col(position = "dodge") +
    geom_text(aes(label = paste(value, "%")), vjust = -1, position = position_dodge(0.9)) +
    scale_y_continuous(limits = c(0, 100)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          text = element_text(size = 16)) +
    labs(
      title = "Neighbourhood Rating in 2019",
      x = "Neighbourhood Rating",
      y = "Percentage of survey responses",
      fill = str_wrap(str_to_title(str_replace_all(deparse(substitute(category)),
                                                   "_", " ")), 20)
    )
}
```

```{r}
#categories <- c("ethnicity", "type_of_tenure", "simd_quintiles", "household_type",
# "urban_rural_classification", "gender", "walking_distance_to_nearest_greenspace")

create_2019_neighbourhood_graph(ethnicity)
create_2019_neighbourhood_graph(type_of_tenure)
create_2019_neighbourhood_graph(simd_quintiles)
create_2019_neighbourhood_graph(household_type)
create_2019_neighbourhood_graph(urban_rural_classification)
create_2019_neighbourhood_graph(gender)
create_2019_neighbourhood_graph(walking_distance_to_nearest_greenspace)
#ggsave("graphs_and_images/neighbourhood_and_greenspace.png")
```

```{r}
neighbourhood_rating %>% 
  filter(area == "Scotland", measurement == "Percent", gender != "All",
         year == "2019") %>% 
  select(gender, neighbourhood_rating, value) %>% 
  ggplot(aes(neighbourhood_rating, value, fill = gender)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = paste(value, "%")), vjust = -1, position = position_dodge(0.9)) +
  scale_y_continuous(limits = c(0, 100)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        text = element_text(size = 16)) +
  labs(
    title = "Neighbourhood Rating in 2019",
    x = "Neighbourhood Rating",
    y = "Percentage of survey responses",
    fill = deparse(substitute(category))
  )
```

```{r}
neighbourhood_rating %>% 
      filter(area == "Scotland", measurement == "Percent",
             walking_distance_to_nearest_greenspace != "All") %>% 
      summarise(mean_score = sum(score),
                .by = c(year, input$variable_input_n)) %>% 
      ggplot(aes(year, mean_score, colour = get(input$variable_input_n))) +
      geom_line() +
      geom_point(size = 5) +
      scale_x_continuous(breaks = seq(2013, 2019, 1)) +
      #scale_y_continuous(limits = c(-1, 1)) +
      geom_vline(xintercept = input$year_input_n,
                 colour = "black", alpha = 0.5) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            text = element_text(size = 24)) +
      labs(
        title = "Neighbourhood Rating Score",
        x = "Year",
        y = "Average Score (-1 to 1)",
        colour = "Classification"
      )
```


```{r}
neighbourhood_rating %>% 
  filter(gender != "All", area == "Scotland", measurement == "Percent", year == 2019) %>% 
```

# Area exploration
```{r}
area_ratings <- neighbourhood_rating %>% 
  filter(measurement == "Percent", gender != "All") %>% 
  group_by(feature_code, area, year, gender) %>% 
  summarise(total_score = sum(score)) %>% 
  ungroup()
```
```{r}
area_ratings_year <- area_ratings %>% 
  group_by(feature_code, area, year) %>% 
  summarise(avg_score = mean(total_score))
```
```{r}
area_ratings_year
```

```{r}
area_ratings_all <- area_ratings %>% 
  group_by(feature_code) %>% 
  summarise(score = mean(total_score))
area_ratings_all
```

## For each year what was the highest rated area?
```{r}
area_ratings %>% 
  group_by(year) %>% 
  slice_max(avg_score, n = 5) %>% 
  ggplot(aes(year, avg_score, fill = area)) +
  geom_col(position = "dodge", show.legend = FALSE) +
  geom_text(aes(label = area, y = 0), position = position_dodge(0.9),
            hjust = "left", size = 3) +
  scale_x_continuous(breaks = seq(2013, 2019, 1)) +
  coord_flip()
```

## Difference in each area since 2013?
```{r}
change_in_score <- area_ratings_year %>% 
  filter(year %in% c(2013, 2019)) %>% 
  group_by(feature_code) %>% 
  summarise(diff = diff(avg_score)) %>%
  arrange(diff)

change_in_score
```

```{r}
library(sf)
```

[Data source:](https://data.spatialhub.scot/dataset/local_authority_boundaries-is/resource/d24c5735-0f1c-4819-a6bd-dbfeb93bd8e4)
```{r}
uk_boundaries <- st_read(
  dsn = here("clean_data/map_data/"),
             layer = "pub_las")
```
```{r}
plot(uk_boundaries)
```

```{r}
names(uk_boundaries)
```
```{r}
head(uk_boundaries)
```

```{r}
joined_areas <- uk_boundaries %>% 
  #select(CODE, geometry) %>% 
  inner_join(area_ratings_all, by = join_by(code == feature_code))
```

```{r}
joined_areas
```
```{r}
plot(joined_areas[4])
```
```{r}
ggplot(joined_areas, aes(fill = score)) +
  geom_sf() +
  scale_fill_distiller(palette = "3-class Greens", direction = 1) +
  theme_minimal() +
  theme(axis.text = element_blank())
```
## Exploring maps for different gender
```{r}
gender_and_area <- neighbourhood_rating %>% 
  filter(measurement == "Percent", gender != "All") %>% 
  group_by(feature_code, area, year, gender) %>% 
  summarise(total_score = sum(score)) %>% 
  group_by(feature_code, area, gender) %>% 
  summarise(score = mean(total_score))
```
```{r}
joined_gender <- uk_boundaries %>% 
  right_join(gender_and_area, by = join_by(code == feature_code))
```

```{r}
ggplot(joined_gender, aes(fill = score)) +
  geom_sf() +
  facet_wrap(~gender)
```
### both together
```{r}
council_boundaries <- st_read(
  dsn = here("clean_data/map_data/"),
  layer = "pub_las")

spatial_neighbourhood <- read_csv(here("clean_data/spatial_neighbourhood.csv")) 

# spatial_neighbourhood_joined <- council_boundaries %>% 
#    right_join(spatial_neighbourhood, by = join_by(code == feature_code))

spatial_community <- read_csv(here("clean_data/spatial_community.csv"))

# spatial_community_joined <- council_boundaries %>% 
#    right_join(spatial_community, by = join_by(code == feature_code))
```
```{r}
spatial_neighbourhood <- spatial_neighbourhood %>% 
  rename("neighbourhood_score" = score) %>% 
  rename("neighbourhood_change" = diff)

spatial_community <- spatial_community %>% 
  rename("community_score" = score) %>% 
  rename("community_change" = diff)

all_joined <- spatial_community %>% 
  full_join(spatial_neighbourhood) %>% 
  pivot_longer(c(community_score, neighbourhood_score),
               names_to = "metric", values_to = "score") %>% 
  mutate(metric = str_remove(metric, "_[a-z]+$")) %>% 
  pivot_longer(c(community_change, neighbourhood_change),
               names_to = "change_metric", values_to = "change") %>% 
  mutate(change_metric = str_remove(change_metric, "_[a-z]+$")) %>% 
  filter(metric == change_metric) %>% 
  select(-change_metric)
```

```{r}
all_joined_spatial <- council_boundaries %>% 
  right_join(all_joined, by = join_by(code == feature_code))
```

```{r}
ggplot(all_joined_spatial, aes(fill = score)) +
  geom_sf() +
  facet_wrap(~metric)
```

# function to make map?
```{r}
make_map <- function(category, fill_type = score){
  neighbourhood <- neighbourhood_rating %>% 
    filter(measurement == "Percent", {{category}} != "All") %>% 
    group_by(feature_code, area, year, {{category}}) %>% 
    summarise(score = sum(score)) %>% 
    group_by(feature_code, area, {{category}}) %>% 
    summarise(score = mean(score))
  
  joined <- council_boundaries %>% 
    right_join(neighbourhood, by = join_by(code == feature_code))
  
  return(joined)
}
```

```{r}
make_map(urban_rural_classification) %>% 
    ggplot(aes(fill = score)) +
    geom_sf () +
    facet_grid(~urban_rural_classification)
```

```{r}
neighbourhood <- neighbourhood_rating %>% 
    filter(measurement == "Percent", gender != "All") %>% 
    group_by(feature_code, area, year, gender) %>% 
    summarise(score = sum(score)) %>% 
    group_by(feature_code, area, gender) %>% 
    summarise(neighbourhood_score = mean(score)) %>% 
    mutate(id = str_c(feature_code, gender)) 

  community <- community_belonging %>% 
    filter(measurement == "Percent", gender != "All") %>% 
    group_by(feature_code, area, year, gender) %>% 
    summarise(score = sum(score)) %>% 
    group_by(feature_code, area, gender) %>% 
    summarise(community_score = mean(score)) %>% 
    mutate(id = str_c(feature_code, gender))
  
  ratings <- neighbourhood %>% 
    select(id, neighbourhood_score) %>% 
    full_join(community, by = join_by(id)) %>% 
    pivot_longer(c(neighbourhood_score, community_score),
                 names_to = "rating_type", values_to = "score") %>% 
    mutate(rating_type = str_remove(rating_type, "_score")) %>% 
    select(feature_code.x, area.x, gender, rating_type, score)
  
  joined <- council_boundaries %>% 
    right_join(ratings, by = join_by(code == feature_code.x))
  
  joined %>% 
    ggplot(aes(fill = score)) +
    geom_sf () +
    facet_grid(rating_type~gender)
```

