---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(here)
```

```{r}
community_belonging <- clean_names(read_csv(here("raw_data/community_belonging.csv")))
green_spaces <- clean_names(read_csv(here("raw_data/green_spaces.csv")))
neighbourhood_rating <- clean_names(read_csv(here("raw_data/neighbourhood_rating.csv")))
```

```{r}
community_belonging
green_spaces
neighbourhood_rating
```
# Explore Community Belonging
```{r}
community_belonging %>% 
  summarise(across(.cols = everything(), .fns = ~sum(is.na(.x))))
```
There are no NA values

```{r}
community_belonging %>% 
  distinct(units)
```
There is only 1 value in this column. It can be removed in the cleaning script

```{r}
community_belonging %>% 
  distinct() %>% 
  nrow()

community_belonging %>% 
  nrow()
```
No duplicates

```{r}
community_belonging %>% 
  filter(feature_code == "S12000036", year == "2016",
         walking_distance_to_nearest_greenspace == "More than 10 minutes")
```
There appear to be some missing values such as how many people voted "Don't know" to community belonging where the walking distance is greater than 10 minutes. Is it sensible to assume this is 0?

```{r}
community_belonging %>% 
  filter(feature_code == "S12000036", year == "2016",
         walking_distance_to_nearest_greenspace == "More than 10 minutes",
         measurement == "Percent") %>% 
  summarise(sum(value))
```
Yes, the others add up to 100% so 0% voted "Don't know".


## What is the granularity of the data? 
```{r}
community_belonging %>% 
  filter(date_code == 2019,
         gender == "Female")
```
The feature codes look like council areas. The name of the area if going to be more useful than codes.
![data source](https://www.opendata.nhs.scot/km/dataset/geography-codes-and-labels)
```{r}
council_areas <- clean_names(read_csv(here("raw_data/council_area_codes.csv")))

council_areas_clean <- council_areas %>% 
  select(ca, ca_name) %>% 
  add_row(ca = "S92000003" ,ca_name = "Scotland")
```
Some rows appear to be for the whole of Scotland so I have added a row for this code.

```{r}
community_belonging <- community_belonging %>% 
  left_join(council_areas_clean, by = c("feature_code" = "ca"))
```

When plotting the order of community belonging will not make sense, therefore this should be factored.
```{r}
community_belonging <- community_belonging %>% 
  mutate(community_belonging = factor(community_belonging, 
                                      levels = c("Not at all strongly", 
                                                 "Not very strongly",
                                                 "Don't know",
                                                 "Fairly strongly",
                                                 "Very strongly")))
```
> Each row appears to be a demographic (eg. Female, more than 10 minutes to nearest greenspace, etc), category of community belonging, measurement, and the percentage of people of that demographic that scored their community belonging at that category. 

```{r}
community_belonging %>% 
  distinct(walking_distance_to_nearest_greenspace)
```

## Data

* feature_code = council area or all of scotland
* date_code = year between 2013 and 2019
* measurement = whether the value is a percent, or confidence limit
* value = percent of people that gave that rating or upper and lower confidence levels for that percentage
* **community_belonging** = rating of community belonging
* gender = Male/Female
* urban_rural_classification = Urban/Rural
* simd_quintiles = 80% least deprived/20% most deprived
* type_of_tenure = Owned Mortgage/Loan / Owned Outright / Social Rented / Private Rented / Other
* household_type = With Children / Pensioners / Adults
* ethnicity = White / Other
* **walking_distance_to_nearest_greenspace** = More than 10 minutes / Less than 10 minutes / Don't know

### Quality of data
```{r}
community_belonging %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(total = n())
```
```{r}
(105/(4107+1035))*100
```
Walking distance to nearest greenspace is likely self reported, only 2% of people do not know the answer and there is a reasonable amount of data that do know.

## Experimental plotting

```{r}
community_belonging %>% 
  filter(date_code == 2019,
         gender == "Female",
         ca_name == "Scotland") %>% 
  select(measurement:community_belonging) %>% 
  pivot_wider(names_from = measurement, values_from = value) %>% 
  clean_names() %>% 
  arrange(community_belonging) %>% 
  ggplot(aes(community_belonging, percent)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = x95_percent_lower_confidence_limit_percent,
                    ymax = x95_percent_upper_confidence_limit_percent,
                    width = 0.2))
```
Does this add up to 100%?
```{r}
community_belonging %>% 
  filter(date_code == 2019,
         gender == "Female",
         ca_name == "West Lothian") %>% 
  select(measurement:community_belonging) %>% 
  pivot_wider(names_from = measurement, values_from = value) %>% 
  clean_names() %>% 
  summarise(sum(percent))
```
Yes.

## Community belonging and green space

```{r}
community_belonging %>% 
  filter(walking_distance_to_nearest_greenspace != "All",
         ca_name == "Scotland",
         date_code == 2019) %>% 
  pivot_wider(names_from = measurement, values_from = value) %>% 
  clean_names() %>% 
  ggplot(aes(community_belonging, percent, fill = walking_distance_to_nearest_greenspace)) +
  geom_col(position = "dodge") +
  facet_wrap(~walking_distance_to_nearest_greenspace)
```
```{r}
community_belonging %>% 
  filter(walking_distance_to_nearest_greenspace != "All",
         ca_name == "Scotland",
         date_code == 2019) %>% 
  pivot_wider(names_from = measurement, values_from = value) %>% 
  clean_names() %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(total = sum(percent))
```
It is difficult to quanity the difference.   
Perhaps we could use a score for community belonging? (0-5 or (-2)-(+2))

```{r}
community_belonging %>% 
  distinct(community_belonging)
```

```{r}
community_belonging_scored <- community_belonging %>% 
  mutate(belonging_score = case_when(
    community_belonging == "Not at all strongly" ~ 1,
    community_belonging == "Not very strongly" ~ 2,
    community_belonging == "Don't know" ~ 3,
    community_belonging == "Fairly strongly" ~ 4,
    community_belonging == "Very strongly" ~ 5,
  ),
  belonging_score_zeroed = case_when(
    community_belonging == "Not at all strongly" ~ -2,
    community_belonging == "Not very strongly" ~ -1,
    community_belonging == "Don't know" ~ 0,
    community_belonging == "Fairly strongly" ~ 1,
    community_belonging == "Very strongly" ~ 2,
  ))
```

```{r}
community_belonging_scored %>% 
  filter(walking_distance_to_nearest_greenspace != "All",
         ca_name == "Scotland",
         date_code == 2019) %>% 
  pivot_wider(names_from = measurement, values_from = value) %>% 
  clean_names() %>% 
  mutate(score = percent * belonging_score,
         score_zero = percent * belonging_score_zeroed) %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(overall_belonging = mean(score),
            overall_belonging_zero = mean(score_zero)) %>% 
  ggplot(aes(walking_distance_to_nearest_greenspace, overall_belonging_zero)) +
  geom_col()
```
With normal scoring the maximum possible is 100% scoring 5 -> 500 (0:500)  
With zeroed score the maximum possible is 100% scoring 2 -> 200 (-200:200)  
This might be better on a scale of -1 to 1.  

```{r}
community_belonging_scored %>% 
  filter(walking_distance_to_nearest_greenspace != "All",
         ca_name == "Scotland",
         date_code == 2019) %>% 
  pivot_wider(names_from = measurement, values_from = value) %>% 
  clean_names() %>% 
  mutate(score_zero = (percent * belonging_score_zeroed)/200) %>% 
  group_by(walking_distance_to_nearest_greenspace) %>% 
  summarise(overall_belonging_zero = mean(score_zero)) %>% 
  ggplot(aes(walking_distance_to_nearest_greenspace, overall_belonging_zero)) +
  geom_col()
``` 

```{r}
community_belonging_scored %>% 
  filter(walking_distance_to_nearest_greenspace != "All",
         ca_name == "Scotland",
         date_code == 2019) %>% 
  pivot_wider(names_from = measurement, values_from = value) %>% 
  clean_names() %>% 
  mutate(score_zero = (percent * belonging_score_zeroed)/200) %>% 
  ggplot(aes(walking_distance_to_nearest_greenspace, score_zero)) +
  geom_point()
```  

App should explore community belonging for each group.

# Green spaces
```{r}
green_spaces
green_spaces %>% 
  distinct(age)
```
* feature_code = council area or all of scotland
* date_code = year between 2013 and 2019
* measurement = whether the value is a percent, or confidence limit
* value = percent of people that gave that rating or upper and lower confidence levels for that percentage
* **distance_to_nearest_green_or_blue_space** = 5 minute walk or less / 6-10 minute walk / 11+ minute walk / Don't know
* age = 16-34 / 35-64 / 65+
* gender = Male/Female
* urban_rural_classification = Urban/Rural
* simd_quintiles = 80% least deprived/20% most deprived
* type_of_tenure = Owned Mortgage/Loan / Owned Outright / Social Rented / Private Rented / Other
* household_type = With Children / Pensioners / Adults
* ethnicity = White / Other

This dataset tells us who has access to green or blue spaces within 10 minutes.

```{r}
green_spaces %>% 
  left_join(council_areas_clean, by = c("feature_code" = "ca")) %>% 
  filter(ca_name == "Scotland",
         urban_rural_classification == "Urban",
         measurement == "Percent") %>% 
  ggplot(aes(date_code, value, colour = distance_to_nearest_green_or_blue_space)) +
  geom_line()
```
# Neighbourhood Rating

```{r}
neighbourhood_rating
```
```{r}
neighbourhood_rating %>% 
  distinct() %>% 
  nrow()

neighbourhood_rating %>% 
  nrow()
```



```{r}
neighbourhood_rating %>% 
  distinct(neighbourhood_rating)
```

# Tab data

```{r}
full_data_2019 <- read.delim(here("raw_data/tab/shs2019_social_public.tab"))
```

```{r}
data_small <- full_data_2019 %>% 
  filter(rand_ok == 1) %>% 
  select(council, # council area
         randage, # age
         randgender, # gender
         RANDETH2012, # ethnicity
         #randecon,
         tenure_harm, #tenure_type
         hhtype_new, # household type (1, 2, 6, 7 = adult, 3, 4, 5 = children, 8 = pensioner)
         greenfar13, # how far to nearest green or blue space
         MD20QUIN, # SIMD (1 = most, 5 = least)
         SHS_2CLA # urban_rural
         #rb1:asb1i
  )
data_small
```
```{r}
data_clean <- data_small %>% 
  rename("ethnicity" = "RANDETH2012") %>% 
  rename("simd" = "MD20QUIN") %>% 
  rename("urban_rural" = "SHS_2CLA") %>% 
  mutate(
    household_type = case_when(
      hhtype_new %in% c(1, 2, 6, 7) ~ "Adult",
      hhtype_new %in% c(3, 4, 5) ~ "Children",
      hhtype_new == 8 ~ "Pensioner"
    ),
    gender = case_when(
      randgender == 1 ~ "Male",
      randgender == 2 ~ "Female",
      .default = "Other/Refused"
    ),
    ethnicity = case_when(
      ethnicity == 1 ~ "White",
      ethnicity == 2 ~ "Other",
      .default = "Refused/Unknown"
    ),
    tenure = case_when(
      tenure_harm == 1 ~ "Owned Outright",
      tenure_harm == 2 ~ "Owned Loan/Mortgage",
      tenure_harm == 3 | tenure_harm == 4  ~ "Social Rented",
      tenure_harm == 5 ~ "Private Rented",
      tenure_harm == 6 ~ "Other",
      .default = "Refused/Unknown"
    ),
    greenspace = case_when(
      greenfar13 == 1 ~ "5 mins or less",
      greenfar13 == 2 ~ "6-10 mins",
      greenfar13 == 3 ~ "11-20 mins",
      greenfar13 == 4 ~ "21-30 mins",
      greenfar13 == 5 ~ "More than 30",
      greenfar13 == 6 ~ "Don't know"
    ),
    urban_rural = case_when(
      urban_rural == 1 ~ "Urban",
      urban_rural == 2 ~ "Rural"
    )
    ) %>% 
  select(-c(hhtype_new, randgender, tenure_harm, greenfar13))
```

```{r}
data_clean %>% 
  summarise(across(.cols = everything(), .fns = ~sum(is.na(.x))))
```
```{r}
data_clean %>% 
  group_by(ethnicity) %>% 
  summarise(n(),
            prop = (n()/nrow(data_clean))*100)
```
```{r}
data_clean %>% 
  group_by(simd) %>% 
  summarise(n(),
            prop = (n()/nrow(data_clean))*100)
```
```{r}
data_clean %>% 
  group_by(urban_rural) %>% 
  summarise(n(),
            prop = (n()/nrow(data_clean))*100)
```
```{r}
data_clean %>% 
  group_by(household_type) %>% 
  summarise(n(),
            prop = (n()/nrow(data_clean))*100)
```
```{r}
data_clean %>% 
  group_by(tenure) %>% 
  summarise(n(),
            prop = (n()/nrow(data_clean))*100)
```
```{r}
data_clean %>% 
  group_by(greenspace) %>% 
  summarise(n(),
            prop = (n()/nrow(data_clean))*100)
```


```{r}
full_data_2019 %>% 
  filter(rand_ok == 1) %>% 
  select(dyear, council, randage, randgender, randecon,
         tenure_harm, hhtype_new, rb1:asb1i) %>% 
  mutate(council = case_when(
    council == "T" ~ "Moray",
    council == "5" ~ "West Lothian",
    council == "Y" ~ "Renfrewshire",
    council == "C" ~ "Angus",
    council == "R" ~ "Inverclyde",
    council == "U" ~ "North Ayrshire",
    council == "X" ~ "Perth and Kinross",
    council == "6" ~ "Eilean Siar",
    council == "B" ~ "Aberdeenshire",
    council == "D" ~ "Argyll and Bute",
    council == "E" ~ "Scottish Borders",
    council == "H" ~ "Dundee City",
    council == "1" ~ "South Ayrshire",
    council == "2" ~ "South Lanarkshire",
    council == "F" ~ "Clackmannanshire",
    council == "G" ~ "Dumfries and Galloway",
    council == "" ~ "",
    council == "" ~ "",
    council == "" ~ "",
    Value =   	Label = 
      Value =   	Label = 
      Value =   	Label = 
      Value = Q  	Label = Highland
    Value = 3  	Label = Stirling
    Value = A  	Label = Aberdeen City
    Value = 4  	Label = West Dumbartonshire
    Value = L  	Label = East Renfrewshire
    Value = N  	Label = Falkirk
    Value = K  	Label = East Lothian
    Value = O  	Label = Fife
    Value = J  	Label = East Dumbartonshire
    Value = W  	Label = Orkney
    Value = P  	Label = Glasgow City
    Value = Z  	Label = Shetland
    Value = V  	Label = North Lanarkshire
    Value = I  	Label = East Ayrshire
    Value = S  	Label = Midlothian
    Value = M  	Label = Edinburgh City
    
  ))
```



























