---
title: "Report on Community Belonging and Neighbourhood Ratings"
author: "Emma Menzies"
output: html_notebook
---

# Libraries and Data

### Libraries
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(infer)
library(GGally)
```

### Data
```{r}
community_belonging <- read_csv(here("clean_data/community_belonging.csv")) %>% 
  mutate(community_belonging = factor(community_belonging, 
                                      levels = c("Don't know",
                                                 "Not at all strongly", 
                                                 "Not very strongly",
                                                 "Fairly strongly",
                                                 "Very strongly")))

neighbourhood_rating <- read_csv(here("clean_data/neighbourhood_rating.csv")) %>% 
  mutate(neighbourhood_rating = factor(neighbourhood_rating, 
                                      levels = c("No opinion",
                                                 "Very poor", 
                                                 "Fairly poor",
                                                 "Fairly good",
                                                 "Very good")))
```

### Functions
```{r}
source(here("analysis_scripts_and_functions/test_diff_in_means.R"))
source(here("analysis_scripts_and_functions/visualise_diff_in_means.R"))
source(here("analysis_scripts_and_functions/create_2019_graph.R"))
```

# 1. Background

## Scottish Household Survey

Every year the ![Scottish Household Survey](https://www.gov.scot/collections/scottish-household-survey/) asks a random selection of adults in private accommodation in Scotland questions about themselves and their household. This survey is run by Ipsos MORI on behalf of the Scottish Government.

This analysis and report will explore the responses given in this survey about neighbourhood ratings and community belonging, and how these are related to other factors such as access to green space, household type, and council area. The data available is from 2013-2019. Due to the COVID pandemic in 2020 the format of the survey had to change to telephone interview. Data from 2020 and 2021 has therefore been explored separately. 

Data Sources:

* ![2013-2019](https://statistics.gov.scot/resource?uri=http%3A%2F%2Fstatistics.gov.scot%2Fdata%2Fcommunity-belonging---shs)
* ![2020](https://www.gov.scot/publications/scottish-household-survey-2020-telephone-survey-key-findings/documents/)
* ![2021](https://www.gov.scot/publications/scottish-household-survey-2021-telephone-survey-key-findings/documents/)

## Ethics

Since the results come from a random selection of the general population they may not be entirely representative. People that are not in private accommodation (eg. prisons, hospitals, student accommodation) will not be represented. Neither will those who refused to participate in the survey. This could lead to some bias in the results.

Since the data collected in the survey contains sensitive information, the data publicly available is highly aggregated. This limits the analysis which can be done, but ensures there is no risk of sensitive information of being shared within this report.

# 2. Data Cleaning and Exploration

* scoring system
* aggregation
* ethics

# 3. Does access to greenspace impact neighbourhood ratings or community belonging?

## Neighbourhood rating

The question asked by the survey is as follows:
>"THINKING NOW ABOUT THE NEIGHBOURHOOD YOU LIVE IN, HOW
WOULD YOU RATE IT AS A PLACE TO LIVE?
If pressed, define 'your neighbourhood' as: 'the street you live in and the
streets nearby' (urban) or 'the local area' (rural)."

The options given for the response were:
>"(1) Very good 
  (2) Fairly good 
  (3) Fairly poor 
  (4) Very poor 
  No opinion"

The first factor to explore is walking distance to greenspace, given in greater than, or less than, 10 minutes.

The percentages indicate the percentage of people from the given group which responded in that way to the survey,
  eg. percentage of people with a walking distance to greenspace less than 10 minutes, that rated their neighbourhood "Very good".

```{r}
create_2019_graph_neighbourhood(walking_distance_to_nearest_greenspace)
```

The percentages are fairly similar across all responses suggesting that access to greenspace does not have a significant effect on neighbourhood ratings. There are some small differences however, those that had access to green space within less than 10 minutes of walking appear to be slightly more likely to rate their neighbourhood "Very good" compared to the other groups. People who do not know how close their nearest greenspace is were more liekly than the other groups to have no opinion on the neighbourhood, or rate it "Fairly poor".

This means.....

## Community Belonging

HOW STRONGLY DO YOU FEEL YOU BELONG TO YOUR IMMEDIATE
NEIGHBOURHOOD?
(1) Very strongly
(2) Fairly strongly
(3) Not very strongly
(4) Not at all strongly
Don't know {Spontaneous] 

```{r}
create_2019_graph_community(walking_distance_to_nearest_greenspace)
```

