---
title: "Report on Community Belonging and Neighbourhood Ratings"
author: "Emma Menzies"
output: html_notebook
---

# Libraries and Data

### Libraries
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(infer)
library(GGally)
```

### Data
```{r}
community_belonging <- read_csv(here("clean_data/community_belonging.csv")) %>% 
  mutate(community_belonging = factor(community_belonging, 
                                      levels = c("Not at all strongly", 
                                                 "Not very strongly",
                                                 "Don't know",
                                                 "Fairly strongly",
                                                 "Very strongly")))
neighbourhood_rating <- read_csv(here("clean_data/neighbourhood_rating.csv"))
```

### Functions
```{r}
source(here("analysis_scripts_and_functions/test_diff_in_means.R"))
source(here("analysis_scripts_and_functions/visualise_diff_in_means.R"))
```

# Background

THINKING NOW ABOUT THE NEIGHBOURHOOD YOU LIVE IN, HOW
WOULD YOU RATE IT AS A PLACE TO LIVE?
If pressed, define 'your neighbourhood' as: 'the street you live in and the
streets nearby' (urban) or 'the local area' (rural).

(1) Very good [1]
(2) Fairly good [2]
(3) Fairly poor [3]
(4) Very poor [4]
No opinion [Spontaneous] [5]

HOW STRONGLY DO YOU FEEL YOU BELONG TO YOUR IMMEDIATE
NEIGHBOURHOOD?
(1) Very strongly
(2) Fairly strongly
(3) Not very strongly
(4) Not at all strongly
Don't know {Spontaneous] 

```{r}
community_belonging %>% 
  filter(area == "Scotland", measurement == "Percent", simd_quintiles != "All",
         year == "2019") %>% 
  select(year, simd_quintiles, community_belonging, value) %>% 
  ggplot(aes(community_belonging, value, fill = simd_quintiles)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = paste(value, "%")), vjust = -1, position = position_dodge(0.9)) +
  #facet_wrap(~simd_quintiles) +
  scale_y_continuous(limits = c(0, 50)) +
  #theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        text = element_text(size = 16)) +
  labs(
    title = "Community Belonging by SIMD Quintile in 2019",
    x = "Community Belonging",
    y = "Percentage of survey responses"
  ) +
  scale_fill_manual(values = c("20% most deprived" = "tomato2", "80% least deprived" = "lightgreen"))
```
```{r}
community_belonging %>% 
  filter(area == "Scotland", measurement == "Percent", urban_rural_classification != "All",
         year == "2019") %>% 
  select(year, urban_rural_classification, community_belonging, value) %>% 
  ggplot(aes(community_belonging, value, fill = urban_rural_classification)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = paste(value, "%")), vjust = -1, position = position_dodge(0.9)) +
  #facet_wrap(~urban_rural_classification) +
  scale_y_continuous(limits = c(0, 50)) +
  #theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        text = element_text(size = 16)) +
  labs(
    title = "Community Belonging in Rural and Urban Areas in 2019",
    x = "Community Belonging",
    y = "Percentage of survey responses",
    fill = "Classification"
  ) +
  scale_fill_manual(values = c("Rural" = "darkgreen", "Urban" = "darkgrey"))
```

# Data Cleaning and Exploration

* scoring system
* aggregation
* ethics

# Does access to greenspace impact neighbourhood ratings or community belonging?
```{r}

```

